---
title: "100-comparing-trees"
author: "Mac Campbell"
date: "April 10, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Create a timetree of host taxa
For a subset of taxa, let's create a timetree.
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggtree)

data<-read_tsv("~/github/AAV-GLUE/tabular/locus/epv-locus-data.tsv")
cerco<-data %>% filter(Family=="Cercopithecidae")
species<-cerco %>% select(Species) %>% unique()

write_tsv(species, "data/Cercopithecidae-species-list.tsv")
```

Now I should just be able to upload this to timetree.org and produce a time tree.
```{r, warning=FALSE, message=FALSE}
tree<-read.tree("data/Cercopithecidae-species-list.nwk")
t<-ggtree(tree)

#Making tip ages = 0 
t$data<-t$data %>% mutate(Age=max(x)-x)

t + geom_tiplab() +
  geom_treescale()+
  xlim(0,25)+
  geom_nodelab(aes(label=round(Age,2)), hjust=0)
```

Awesome!!!

## Now, let's get some EVE trees.
There are two goals here.    
__1__ Identify the age of integration      
__2__ Quantifying topological differences between the species tree and the EVE tree    

As for topological differences, I don't know how the rooting will play out. Also, if the EVE is only found in one taxon I imagine we can only go to the divergence of it from the others?? Colobus-angolensis-palliatus-Dependo-85 could have bounds (0, Root Age). So for minimum ages this should be pretty straighforward.

```{r, warning=FALSE, message=FALSE}
library(phytools)

#Making "Taxa" for tip labs and calculate number of taxa
df<-cerco %>% mutate(Delim=gsub(" ","_",Species)) %>%
  group_by(locus_name) %>% mutate(Taxa=paste(Delim, collapse=" "))  %>% mutate(Count=n())

#getting min age of nodes
df2<-df %>% mutate(node = as.numeric(ifelse(Count > 1, findMRCA(tree, tips=strsplit(Taxa, " ")[[1]]), NA)))
df3<-left_join(df2, t$data, by="node")
df3$Age<-replace_na(df3$Age, 0)

#Summarize min age of nodes
df3 %>% select(locus_name, Age) %>% unique()
```

How about quantifying the differences in topology? It may be that we need to prune and unroot our starting time tree.