---
title: "101-full-tree"
author: "Mac Campbell"
date: "April 11, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#How about all the species?
```{r, warning=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggtree)
library(ggrepel)
library(phangorn)
library(ape)
library(mixtools)

data<-read_tsv("~/github/AAV-GLUE/tabular/locus/epv-locus-data.tsv")
species<-data %>% select(Species) %>% unique()
write_tsv(species, "data/species-list.tsv")

```

Queried timetree.org
Four unresolved names:     

Equus caballus fjord      
Equus caballus icelandic      
Neophocaena asiaeorientalis     
     
Twenty-seven species dropped in this approach.
 [1] "Canis familiaris"                     "Canis familiaris"                     "Canis lupus dingo"                   Not in timetree
 
 [4] "Ceratotherium simum cottoni"     no data for Ceratotherium      
  "Cervus elaphus hippelaphus"           "Colobus angolensis palliatus"        Simplified to genus-specie   
 [7] "Dicerorhinus sumatrensis sumatrensis" "Dicerorhinus sumatrensis sumatrensis" "Dicerorhinus sumatrensis sumatrensis"
[10] "Equus africanus asinus"               "Equus africanus asinus"               "Equus asinus somalicus"              
[13] "Equus asinus somalicus"               "Equus burchellii boehmi"              "Equus burchellii boehmi"             
[16] "Equus caballus fjord"                 "Equus caballus icelandic"             "Equus ferus przewalskii"             
[19] "Equus zebra hartmannae"               "Equus zebra hartmannae"               "Murina aurata feae"                  
[22] "Mus castaneus"                        "Mus castaneus"                        "Neomonachus schauinslandi"           
[25] "Neophocaena asiaeorientalis"          "Neophocaena asiaeorientalis"          "Panthera tigris altaica"             

I think I'll edit data file.
```{r, warning=FALSE, message=FALSE}
data<-read_tsv("data/epv-locus-data-for-timetree.tsv")
species<-data %>% select(Species) %>% unique()
write_tsv(species, "data/species-list-curated.tsv")
```

Now we have one unresolved names:    
Neophocaena asiaeorientalis

```{r, warning=FALSE, message=FALSE}
tree<-read.tree("data/species-list-curated.nwk")
t<-ggtree(tree)

#Making tip ages = 0 
t$data<-t$data %>% mutate(Age=max(x)-x)

t + geom_tiplab(size=2) +
  geom_treescale()+
  xlim(0,350)+
  geom_nodelab(aes(label=round(Age,2)), hjust=0, size=2)


ggsave("outputs/101/species-list-timetree.pdf", height=20, width=8.5)

```

That looks ok! 
Compute the minimum ages.
```{r, warning=FALSE, message=FALSE}
library(phytools)
#Making "Taxa" for tip labs and calculate number of taxa
#Exlude unassigned "NK" locus_names
#filter to tips in tree
treetips<-t$data %>% filter(isTip==TRUE) %>% select(label)

df<-data %>% filter(locus_name != "NK") %>%
  mutate(Delim=gsub(" ","_",Species)) %>%
  filter(Delim %in%treetips$label) %>%
  group_by(locus_name) %>% mutate(Taxa=paste(Delim, collapse=" "))  %>% mutate(Count=n())

#getting min age of nodes
df2<-df %>% mutate(node = as.numeric(ifelse(Count > 1, findMRCA(tree, tips=strsplit(Taxa, " ")[[1]]), NA)))
df3<-left_join(df2, t$data, by="node")
df3$Age<-replace_na(df3$Age, 0)

#Summarize min age of nodes
df3 %>% select(locus_name, Age) %>% unique()

annotations<-df3 %>% select(locus_name, node, Age, x, y) %>% unique()
```

# Annotating the tree
We have it set up so that we can select nodes and then also tips based on an "annotations" data frame.

```{r}

nodes<-filter(annotations, node != "NA")
tips<-filter(annotations, Age == 0)
tipSpecies<-df3 %>% filter(locus_name %in% tips$locus_name)

#formatting for carpet plot
carpet<-annotations %>% mutate(NewX=ifelse(Age==0, max(t$data$x), x)) %>% mutate(NewY=ifelse(Age==0, 0, y))
carpet<-carpet %>% mutate(isTip = ifelse(Age==0, "Tip", "Node"))
t + geom_tiplab(size=2) +
  xlim(0,400)+
  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
  geom_tiplab(aes(subset=label %in% tipSpecies$Delim), size=2, color="red")+
  geom_label_repel(data = nodes, aes(x=x, y=y, label=locus_name), size=2, alpha=0.75)+
  geom_rug(data=filter(carpet, isTip=="Node"), aes(x=NewX, y=NewY), sides="b", position="jitter")  +
  theme_tree2()
#max is 312 MYA

ggsave("outputs/101/species-list-min-ages-timetree.pdf", height=20, width=8.5)

```

Lets produce some histograms.
```{r}
carpet2<-carpet %>% ungroup() %>% mutate(mya=max(NewX)-NewX)
ggplot(carpet2) + 
  geom_histogram(aes(x=mya)) + 
  #geom_density(aes(x=mya))+
  facet_grid(.~isTip) + 
  theme_bw()+
  ylab("Count") +
  xlab("Minimum Age of EVE") +
  theme(panel.grid = element_blank())
ggsave("outputs/101/min-eves-histograms.pdf")
```

Also, I can apply a mixture model.
```{r}
internal<-filter(carpet2, mya > 0)

ages = internal$Age
mix = normalmixEM(ages, lambda = 0.75)

summary(mix)
plot(mix,which=2)
lines(density(ages), lty=2, lwd=2)

##ggplot version?
#Test data works ok, but, my real data needs to be bounded somehow..
#https://stackoverflow.com/questions/25313578/any-suggestions-for-how-i-can-plot-mixem-type-data-using-ggplot2
sdnorm = function(x, mean=0, sd=1, lambda=1){lambda*dnorm(x, mean=mean, sd=sd)}

xMix<-mix
dataframe<-as_data_frame(x) %>% rename(x=value)

ggplot(dataframe) + 
 geom_histogram(aes(x=x,y=..density..),fill="white",color="black") +
 stat_function(fun=sdnorm,
    args=list(mean=xMix$mu[2],
             sd=xMix$sigma[2],
             lambda=xMix$lambda[2]),
             fill="#FF000080",geom="polygon") +
 stat_function(fun=sdnorm,
    args=list(mean=xMix$mu[1],
             sd=xMix$sigma[1],
             lambda=xMix$lambda[1]),
             fill="#00FF0080",geom="polygon")+
  xlim(-25,100)

gg.mixEM <- function(EM) {
  require(ggplot2)
  x       <- with(EM,seq(min(x-35),max(x+25),len=1000))
  pars    <- with(EM,data.frame(comp=colnames(posterior), mu, sigma,lambda))
  em.df   <- data.frame(x=rep(x,each=nrow(pars)),pars)
  em.df$y <- with(em.df,lambda*dnorm(x,mean=mu,sd=sigma))
  ggplot(data.frame(x=EM$x),aes(x,y=..density..)) + 
    geom_histogram(fill=NA,color="black")+
    geom_polygon(data=em.df,aes(x,y,fill=comp),color="grey50", alpha=0.5)+
    scale_fill_discrete("Component\nMeans",labels=format(em.df$mu,digits=3))+
    theme_bw()+
    theme(panel.grid = element_blank()
          )+
    ggtitle(label=paste("Mixture with loglik of ", EM$loglik, sep=""))+
    theme(plot.title = element_text(hjust = 0.5))+
    xlab("Age of Endogenization Event")+
    ylab("Density")

}
c2<-mix
c3<-normalmixEM(ages, k=3, lambda = 0.75)
c4<-normalmixEM(ages, k=4, lambda = 0.75)
c5<-normalmixEM(ages, k=5, lambda = 0.75)
c6<-normalmixEM(ages, k=6, lambda = 0.75)

k2<-gg.mixEM(c2)
k3<-gg.mixEM(c3)
k4<-gg.mixEM(c4)
k5<-gg.mixEM(c5)
k6<-gg.mixEM(c6)

library("ggpubr")
ggarrange(k2, k3, k4, k5, k6, ncol = 3, nrow = 2)

ggsave("outputs/101/eve-age-mixtures.pdf", width=22, height=8)
```

## The full tree needs help
__1__ Common names.
__2__ The x-axis.

I'm using the common name files from Shannon. During some curation.
`cat commonnames.tsv | perl -pe "s/_\d+//g" > commonnames-curated.tsv`    
editing subspecific names out    

```{r}
taxonomy<-read_tsv("data/commonnames-curated.tsv") %>% rename(label=MERGE_NAME)
s<-t
s$data <-left_join(s$data,taxonomy)
s$data <-s$data %>% mutate(CommonName=ifelse(COMMON_NAME=="NA", label, COMMON_NAME ))
st<-s + geom_tiplab(aes(label=CommonName),size=2) +
  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
  geom_tiplab(aes(subset=label %in% tipSpecies$Delim, label=CommonName), size=2, color="red")+
  geom_label_repel(data = nodes, aes(x=x, y=y, label=locus_name), size=2, alpha=0.75)+
  theme_tree2()+
  scale_x_continuous(name="Time in Million of Years Ago", breaks=c(12,112,212,312), labels=c(300,200,100,0), limits=c(0,350))

st
ggsave("outputs/101/AAV-timetree.pdf", height=20, width=8.5)

```

```{r}
su<-s+coord_geo(xlim = c(-50,300), ylim=c(0,1),
                dat=list("epochs","periods"),
                pos = as.list(rep("bottom", 2)),
                abbrv = FALSE,
                height = list(unit(4, "lines"), unit(2, "lines")),
                rot = list(90,0),
                size = list(2.5, 5)
                )+scale_x_reverse()

gg<-ggarrange2(st,su, ncol=1, heights = c(19,1))

pdf("outputs/101/AAV-timetree-geologic-timescale.pdf", height=20, width=8.5)
gg
dev.off()

```


 
 