---
title: "101-full-tree"
author: "Mac Campbell"
date: "April 11, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

#How about all the species?
```{r, eval=FALSE}
library(tidyverse)
library(ggtree)
library(ggrepel)
library(phangorn)
library(ape)
library(mixtools)

#data<-read_tsv("~/github/AAV-GLUE/tabular/locus/epv-locus-data.tsv")

#Can't find this file anymore, using a new one, adding in birds and reptile(s), moving to a local spot

#data<-read_tsv("~/github/AAV-GLUE/tabular/epv-digs/epv-side-data.tsv") %>% filter(locus_name !="NK")
data<-read_tsv("data/epv-side-data.tsv")
species<-data %>% select(Species) %>% unique()
write_tsv(species, "data/species-list.tsv", col_names = FALSE)

```

05262020: This file seems to be missing Aves+Reptilia. Adding them in and querying timetree.

17 unresolved names:     
Equus caballus fjord    
Equus caballus icelandic     
Panthera tigris altaica (replaced with Panthera tigris)    
Canis lupus dingo (replaced with Canis lupus)    
Canis familiaris (replaced with Canis lupus)    
Equus asinus somalicus (replaced with Equus asinus)    
Equus africanus asinus (replaced with Equus asinus)    
Equus burchellii boehmi (replaced with Equus burchellii)    
Equus zebra hartmannae (replaced with Equus zebra)    
Ceratotherium simum cottoni (replaced with Ceratotherium simum)    
Dicerorhinus sumatrensis sumatrensis (replaced with Dicerorhinus sumatrensis)    
Cervus elaphus hippelaphus (replaced with Cervus elaphus)    
Murina aurata feae (replaced with Murina aurata)    
Mus castaneus (replaced with Mus musculus)    
Peromyscus maniculatus bairdii (replaced with Peromyscus maniculatus)    
Colobus angolensis palliatus (replaced with Colobus angolensis)    
Neophocaena asiaeorientalis (no substitute found)    

I can edit the species file, previously I did this:
```{r, eval=FALSE}
data<-read_tsv("data/epv-locus-data-for-timetree.tsv")
species<-data %>% select(Species) %>% unique()
write_tsv(species, "data/species-list-curated.tsv", col_names = FALSE)
```

That leaves one unresolved names:    
Neophocaena asiaeorientalis     

But, I think we need a strategy to resolve a couple of things. Perhaps some manual work on the timetree. I've added subspecies epithets for the tiger, red deer and monkey.     

```{r}
#tree<-read.tree("data/species-list-curated.nwk")
tree<-read.tree("data/species-list.nwk")
t<-ggtree(tree)

#Making tip ages = 0 
t$data<-t$data %>% mutate(Age=max(x)-x)

t + geom_tiplab(size=2) +
  geom_treescale()+
  xlim(0,350)+
  geom_nodelab(aes(label=round(Age,2)), hjust=0, size=2)


ggsave("outputs/101/species-list-timetree.pdf", height=20, width=8.5)

```

That looks ok! 
Compute the minimum ages.
```{r, warning=FALSE, message=FALSE}
library(phytools)
#Making "Taxa" for tip labs and calculate number of taxa
#Exlude unassigned "NK" locus_names
#filter to tips in tree
treetips<-t$data %>% filter(isTip==TRUE) %>% select(label)

df<-data %>% filter(locus_name != "NK") %>%
  mutate(Delim=gsub(" ","_",Species)) %>%
  filter(Delim %in% treetips$label) %>%
  group_by(locus_name) %>% mutate(Taxa=paste(Delim, collapse=" "))  %>% mutate(Count=n())

#getting min age of nodes
df2<-df %>% mutate(node = as.numeric(ifelse(Count > 1, findMRCA(tree, tips=strsplit(Taxa, " ")[[1]]), NA)))
df3<-left_join(df2, t$data, by="node")
df3$Age<-replace_na(df3$Age, 0)


#Summarize min age of nodes
df3 %>% select(locus_name, Age) %>% unique()

annotations<-df3 %>% select(locus_name, node, Age, x, y) %>% unique()

```

# Annotating the tree
We have it set up so that we can select nodes and then also tips based on an "annotations" data frame.

```{r}

nodes<-filter(annotations, node != "NA")
tips<-filter(annotations, Age == 0)
tipSpecies<-df3 %>% filter(locus_name %in% tips$locus_name)

#formatting for carpet plot
carpet<-annotations %>% mutate(NewX=ifelse(Age==0, max(t$data$x), x)) %>% mutate(NewY=ifelse(Age==0, 0, y))
carpet<-carpet %>% mutate(isTip = ifelse(Age==0, "Tip", "Node"))
u<-t + geom_tiplab(size=2) +
  xlim(0,400)+
  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
  geom_tiplab(aes(subset=label %in% tipSpecies$Delim), size=2, color="red")+
  geom_label_repel(data = nodes, aes(x=x, y=y, label=locus_name), size=2, alpha=0.75)+
  geom_rug(data=filter(carpet, isTip=="Node"), aes(x=NewX, y=NewY), sides="b", position="jitter")  +
  theme_tree2()+
  scale_x_continuous(name="Time in Million of Years Ago", breaks=c(12,112,212,312),
                     labels=c(300,200,100,0), limits=c(0,350))
#max is 312 MYA
u
ggsave("outputs/101/species-list-min-ages-timetree.pdf", height=20, width=8.5)

```
## What about some clade labels?

We'll need all the metadata.
```{r}
clades<-df %>% select(Delim, Class, Superorder, Order)
orders <- clades %>% ungroup %>% select(Order) %>% unique() %>% arrange(Order)

Carnivora <- filter(clades, Order == "Carnivora")
CarnivoraNode <- findMRCA(tree, tips=Carnivora$Delim, type = "node")

Diprotodontia <- filter(clades, Order =="Diprotodontia")
DiprotodontiaNode <- findMRCA(tree, tips=Diprotodontia$Delim, type = "node")


u + geom_cladelabel(CarnivoraNode, "Carnivora", offset=75) +
    geom_cladelabel(DiprotodontiaNode, "Diprotodontia", offset=75) +
    xlim(0,450)

```

I think I should write a function.

```{r}

#Order is a character string
#Clades is a dataframe "clades" in this markdown
#Tree is our tree, here it is tree
Clades<-clades
Tree<-tree
findMeANode<-function(Taxon) {
  species <- dplyr::filter(Clades, Order == Taxon)
  speciesNode <- findMRCA(Tree, tips=species$Delim)
  return(speciesNode)
}

ovector<-orders$Order
findMeANode("Carnivora")
nodes<-lapply(ovector, findMeANode)

nodes[sapply(nodes, is.null)] <- NA

n<-as_tibble(as.vector(unlist(nodes)))

cladelabels<-bind_cols(orders,n) %>% filter(value != "NA")

```

Looks like I have to incldue th eclade labels one at a time.
```{r}
cladelabels<-filter(cladelabels, !(Order %in% c("Unclassified") ))
                    
nn<-cladelabels$value
ll<-cladelabels$Order
u + 
    geom_cladelabel(nn[1],ll[1], offset=75) +
    geom_cladelabel(nn[2],ll[2], offset=75) +
    geom_cladelabel(nn[3],ll[3], offset=75) +
    geom_cladelabel(nn[4],ll[4], offset=75) +
    geom_cladelabel(nn[5],ll[5], offset=75) +
    geom_cladelabel(nn[6],ll[6], offset=75) +
#    geom_cladelabel(nn[7],ll[7], offset=75) +
    geom_cladelabel(nn[8],ll[8], offset=75) +
    geom_cladelabel(nn[9],ll[9], offset=75) +
    geom_cladelabel(nn[10],ll[10], offset=75) +
    geom_cladelabel(nn[11],ll[11], offset=75) +
    geom_cladelabel(nn[12],ll[12], offset=75) +
    xlim(0,450)


ggsave("outputs/101/species-list-min-ages-timetree-clades.pdf", height=20, width=9)

```

Lets produce some histograms.
```{r}
carpet2<-carpet %>% ungroup() %>% mutate(mya=max(NewX)-NewX)
ggplot(carpet2) + 
  geom_histogram(aes(x=mya)) + 
  #geom_density(aes(x=mya))+
  facet_grid(.~isTip) + 
  theme_bw()+
  ylab("Count") +
  xlab("Minimum Age of EVE") +
  theme(panel.grid = element_blank())
ggsave("outputs/101/min-eves-histograms.pdf")
```

Also, I can apply a mixture model.
```{r}
internal<-filter(carpet2, mya > 0)

ages = internal$Age
mix = normalmixEM(ages, lambda = 0.75)

summary(mix)
plot(mix,which=2)
lines(density(ages), lty=2, lwd=2)

##ggplot version?
#Test data works ok, but, my real data needs to be bounded somehow..
#https://stackoverflow.com/questions/25313578/any-suggestions-for-how-i-can-plot-mixem-type-data-using-ggplot2
sdnorm = function(x, mean=0, sd=1, lambda=1){lambda*dnorm(x, mean=mean, sd=sd)}

xMix<-mix
dataframe<-as_data_frame(x) %>% rename(x=value)

ggplot(dataframe) + 
 geom_histogram(aes(x=x,y=..density..),fill="white",color="black") +
 stat_function(fun=sdnorm,
    args=list(mean=xMix$mu[2],
             sd=xMix$sigma[2],
             lambda=xMix$lambda[2]),
             fill="#FF000080",geom="polygon") +
 stat_function(fun=sdnorm,
    args=list(mean=xMix$mu[1],
             sd=xMix$sigma[1],
             lambda=xMix$lambda[1]),
             fill="#00FF0080",geom="polygon")+
  xlim(-25,100)

gg.mixEM <- function(EM) {
  require(ggplot2)
  x       <- with(EM,seq(min(x-35),max(x+25),len=1000))
  pars    <- with(EM,data.frame(comp=colnames(posterior), mu, sigma,lambda))
  em.df   <- data.frame(x=rep(x,each=nrow(pars)),pars)
  em.df$y <- with(em.df,lambda*dnorm(x,mean=mu,sd=sigma))
  ggplot(data.frame(x=EM$x),aes(x,y=..density..)) + 
    geom_histogram(fill=NA,color="black")+
    geom_polygon(data=em.df,aes(x,y,fill=comp),color="grey50", alpha=0.5)+
    scale_fill_discrete("Component\nMeans",labels=format(em.df$mu,digits=3))+
    theme_bw()+
    theme(panel.grid = element_blank()
          )+
    ggtitle(label=paste("Mixture with loglik of ", EM$loglik, sep=""))+
    theme(plot.title = element_text(hjust = 0.5))+
    xlab("Age of Endogenization Event")+
    ylab("Density")

}
c2<-mix
c3<-normalmixEM(ages, k=3, lambda = 0.75)
c4<-normalmixEM(ages, k=4, lambda = 0.75)
c5<-normalmixEM(ages, k=5, lambda = 0.75)
c6<-normalmixEM(ages, k=6, lambda = 0.75)

k2<-gg.mixEM(c2)
k3<-gg.mixEM(c3)
k4<-gg.mixEM(c4)
k5<-gg.mixEM(c5)
k6<-gg.mixEM(c6)

library("ggpubr")
ggarrange(k2, k3, k4, k5, k6, ncol = 3, nrow = 2)

ggsave("outputs/101/eve-age-mixtures.pdf", width=22, height=8)
```

## The full tree needs help
__1__ Common names.
__2__ The x-axis.

I'm using the common name files from Shannon. During some curation.
`cat commonnames.tsv | perl -pe "s/_\d+//g" > commonnames-curated.tsv`    
editing subspecific names out    

```{r}
taxonomy<-read_tsv("data/commonnames-curated.tsv") %>% rename(label=MERGE_NAME)
s<-t
s$data <-left_join(s$data,taxonomy)
s$data <-s$data %>% mutate(CommonName=ifelse(COMMON_NAME=="NA", label, COMMON_NAME ))
st<-s + geom_tiplab(aes(label=CommonName),size=2) +
  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
  geom_tiplab(aes(subset=label %in% tipSpecies$Delim, label=CommonName), size=2, color="red")+
  geom_label_repel(data = nodes, aes(x=x, y=y, label=locus_name), size=2, alpha=0.75)+
  theme_tree2()+
  scale_x_continuous(name="Time in Million of Years Ago", breaks=c(12,112,212,312), labels=c(300,200,100,0), limits=c(0,350))

st
ggsave("outputs/101/AAV-timetree.pdf", height=20, width=8.5)

```

```{r}
su<-s+coord_geo(xlim = c(-50,300), ylim=c(0,1),
                dat=list("epochs","periods"),
                pos = as.list(rep("bottom", 2)),
                abbrv = FALSE,
                height = list(unit(4, "lines"), unit(2, "lines")),
                rot = list(90,0),
                size = list(2.5, 5)
                )+scale_x_reverse()

gg<-ggarrange2(st,su, ncol=1, heights = c(19,1))

pdf("outputs/101/AAV-timetree-geologic-timescale.pdf", height=20, width=8.5)
gg
dev.off()

```


 
 