---
title: "102-parvoviridae"
author: "Mac Campbell"
date: "October 27, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ggtree)
library(phytools)
library(ggrepel)
```

## Mega time tree
We have several viral genera now.    

```{r}
amdo<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-amdo-side-data.tsv") %>% 
  select(organism, virus_genus, sequenceID,Class) %>% mutate(locus_name=paste0("amdo-",sequenceID))
dependo<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-dependo-side-data.tsv") %>%
    select(organism, sequenceID, locus_name,host_class) %>% mutate(virus_genus="Dependoparvovirus") %>%
  mutate(Class=host_class) %>% select(-host_class)
erythro<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-erythyro-side-data.tsv") %>% 
  select(organism, virus_genus, sequenceID,Class) %>% mutate(locus_name=paste0("amdo-",sequenceID))
proto<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-proto-side-data.tsv")  %>%
    select(organism, virus_genus, sequenceID, locus_name, Class) 

data<-rbind(amdo,dependo,erythro,proto)
data$Species<-gsub("_"," ",data$organism)
data<-data %>% select(Species, organism, locus_name, virus_genus, sequenceID, Class)
species<-data %>% select(Species) %>% unique()
write_tsv(species, "data/species-list-102.tsv", col_names = FALSE)

```

Adding on writing species names to file:
```{r}
list<-split(data, f=data$locus_name)
```

```{r}
writeList<-function(table) {
  species<-as_tibble(unique(table$Species))
  if(length(species$value) > 1) {
  write_tsv(species, col_names = FALSE, paste0("outputs/102/lists/",unique(table$locus_name),".tsv"))
  }
}
```

```{r}
lapply(list, writeList)
```


18 unresolved names: 
Equus caballus icelandic
Equus caballus fjord
Panthera tigris altaica (replaced with Panthera tigris)
Canis familiaris (replaced with Canis lupus)
Canis lupus dingo (replaced with Canis lupus)
Equus asinus somalicus (replaced with Equus asinus)
Equus africanus asinus (replaced with Equus asinus)
Equus burchellii boehmi (replaced with Equus burchellii)
Equus zebra hartmannae (replaced with Equus zebra)
Ceratotherium simum cottoni (replaced with Ceratotherium simum)
Dicerorhinus sumatrensis sumatrensis (replaced with Dicerorhinus sumatrensis)
Giraffa camelopardalis tippelskirchi (replaced with Giraffa camelopardalis)
Cervus elaphus hippelaphus (replaced with Cervus elaphus)
Murina aurata feae (replaced with Murina aurata)
Mus castaneus (replaced with Mus musculus)
Peromyscus maniculatus bairdii (replaced with Peromyscus maniculatus)
Colobus angolensis palliatus (replaced with Colobus angolensis)
Neophocaena asiaeorientalis (no substitute found)

```{r}
tree<-read.tree("data/species-list-102.nwk")
t<-ggtree(tree)

#Making tip ages = 0 
t$data<-t$data %>% mutate(Age=max(x)-x)

t + geom_tiplab(size=2) +
  geom_treescale()+
  xlim(0,500)+
  geom_nodelab(aes(label=round(Age,2)), hjust=0, size=2)


ggsave("outputs/101/species-list-102-timetree.pdf", height=20, width=8.5)

```


```{r, warning=FALSE, message=FALSE}
#Making "Taxa" for tip labs and calculate number of taxa
#Exlude unassigned "NK" locus_names
#filter to tips in tree
treetips<-t$data %>% filter(isTip==TRUE) %>% select(label)

df<-data %>% filter(locus_name != "NK") %>% unique() %>%
  mutate(Delim=gsub(" ","_",Species)) %>%
  filter(Delim %in% treetips$label) %>%
  group_by(locus_name) %>% mutate(Taxa=paste(Delim, collapse=" "))  %>% mutate(Count=n())

#getting min age of nodes
df2<-df %>% mutate(node = as.numeric(ifelse(Count > 1, findMRCA(tree, tips=strsplit(Taxa, " ")[[1]]), NA)))
df3<-left_join(df2, t$data, by="node")
df3$Age<-replace_na(df3$Age, 0)


#Summarize min age of nodes
df3 %>% select(locus_name, Age) %>% unique()

annotations<-df3 %>% select(locus_name, node, Age, x, y, Class, virus_genus) %>% unique()

minAges<-annotations %>% rename(MinimumAge=Age) %>% select(virus_genus, locus_name, MinimumAge, Class)

write_csv(minAges, "outputs/102/parvoviridae-min-ages.csv")
```

##Annotating tree
```{r}

nodes<-filter(annotations, node != "NA")
tips<-filter(annotations, Age == 0)
tipSpecies<-df3 %>% filter(locus_name %in% tips$locus_name)

#formatting for carpet plot
carpet<-annotations %>% mutate(NewX=ifelse(Age==0, max(t$data$x), x)) %>% mutate(NewY=ifelse(Age==0, 0, y))
carpet<-carpet %>% mutate(isTip = ifelse(Age==0, "Tip", "Node"))
u<-t + geom_tiplab(size=2) +
  xlim(0,435)+
  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
  geom_tiplab(aes(subset=label %in% tipSpecies$Delim), size=2, color="grey")+
  geom_label_repel(data = nodes, aes(x=x, y=y, label=locus_name), size=2, alpha=0.75)+
  geom_rug(data=filter(carpet, isTip=="Node"), aes(x=NewX, y=NewY, color=virus_genus), sides="b", position="jitter")  +
  theme_tree2()+
  scale_x_continuous(name="Time in Million of Years Ago", breaks=c(35,135,235,335,435),
                     labels=c(400,300,200,100,0), limits=c(0,550))
u
ggsave("outputs/102/parvoviridae-min-ages-timetree.pdf", height=20, width=8.5)

```