---
title: "105-parvoviridae-tree"
author: "Mac Campbell"
date: "March 12, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ggtree)
library(phytools)
library(ggrepel)
```

## Mega time tree
We have several viral genera now.    

```{r}
amdo<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-amdo-side-data.tsv") %>% 
  select(organism, sequenceID, locus_name, host_class) %>%
  mutate(virus_genus = "Amdoparvovirus") %>% rename(Class=host_class)

dependo<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-dependo-side-data.tsv") %>%
    select(organism, sequenceID, locus_name, host_class) %>% mutate(virus_genus="Dependoparvovirus") %>%
  rename(Class=host_class) 

erythro<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-erythyro-side-data.tsv") %>% 
  select(organism, sequenceID, locus_name, host_class) %>% 
  mutate(virus_genus = "Erythroparvovirus") %>% rename(Class=host_class)

ichthama<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-ichthama-side-data.tsv") %>%
  select(organism, sequenceID, locus_name, host_class) %>%
  mutate(virus_genus = "Ichthamaparvovirus")

proto<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-proto-side-data.tsv")  %>%
  select(host_species, sequenceID, locus_name, host_class) %>%
  rename(organism=host_species) %>% 
  select(organism, sequenceID, locus_name, host_class) %>%
  mutate(virus_genus="Protoparvovirus")

```

Creating a list of species to generate a timetree
```{r}
data<-bind_rows(amdo,dependo,erythro,ichthama, proto)
data$Species<-gsub("_"," ",data$organism)
data<-data %>% select(Species, organism, locus_name, virus_genus, sequenceID, Class)
species<-data %>% select(Species) %>% unique()
write_tsv(species, "outputs/105/species-list-105.tsv", col_names = FALSE)

```

Adding on writing species names to file:
```{r}
list<-split(data, f=data$locus_name)
```

```{r}
writeList<-function(table) {
  species<-as_tibble(unique(table$Species))
  if(length(species$value) > 1) {
  write_tsv(species, col_names = FALSE, paste0("outputs/105/lists/",unique(table$locus_name),".tsv"))
  }
}
```

```{r}
lapply(list, writeList)
```

##Make a time tree

20 unresolved names: 

Lonchura striata domestica (replaced with Lonchura striata) (ok)      
Pelusios castaneus (replaced with Pelusios williamsi) (changed back)      
Panthera tigris altaica (replaced with Panthera tigris) (changed back)     
Prionailurus bengalensis euptilurus (replaced with Prionailurus bengalensis) (changed back)     
Canis lupus dingo (replaced with Canis lupus) (changed both dogs to familiaris)      
Canis familiaris (replaced with Canis lupus)    
Equus africanus asinus (replaced with Equus asinus) (left)    
Equus asinus somalicus (replaced with Equus asinus) (left)     
Equus burchellii boehmi (replaced with Equus burchellii) (changed back) 
Equus zebra hartmannae (replaced with Equus zebra) (changed_back), check > data %>% filter(str_detect(organism, "Equus")) %>% select(Species)
   
Ceratotherium simum cottoni (replaced with Ceratotherium simum), there are both C. s. and C. s. cottoni, left along.   
Dicerorhinus sumatrensis sumatrensis (replaced with Dicerorhinus sumatrensis), changed back.
Cervus elaphus hippelaphus (replaced with Cervus elaphus) (changed back)       
Murina aurata feae (replaced with Murina aurata) (changed back)
Mus castaneus (replaced with Mus musculus) (left alone)
Peromyscus polionotus subgriseus (replaced with Peromyscus polionotus) (changed back)      
Peromyscus maniculatus bairdii (replaced with Peromyscus maniculatus) (chaned back)     
Colobus angolensis palliatus (replaced with Colobus angolensis)
Neophocaena asiaeorientalis (no substitute found), it is basically a porpoise, so adding another tip here.
Emydocephalus ijimae (no substitute found) (in Hydrophiinae, added another tip at 9.2 mya following timetree)   

```{r}
tree<-read.nexus("outputs/105/species-edited.nex")
t<-ggtree(tree)

#Making tip ages = 0 
t$data<-t$data %>% mutate(Age=max(x)-x)

t + geom_tiplab(size=2) +
  geom_treescale()+
  xlim(0,500)+
  geom_nodelab(aes(label=round(Age,2)), hjust=0, size=2)


ggsave("outputs/105/species-list-105-timetree.pdf", height=20, width=8.5)

```


```{r, warning=FALSE, message=FALSE}
#Making "Taxa" for tip labs and calculate number of taxa
#Exlude unassigned "NK" locus_names
#filter to tips in tree
treetips<-t$data %>% filter(isTip==TRUE) %>% select(label)

df<-data %>% filter(locus_name != "NK") %>% unique() %>%
  mutate(Delim=gsub(" ","_",Species)) %>%
  filter(Delim %in% treetips$label) %>%
  group_by(locus_name) %>% mutate(Taxa=paste(Delim, collapse=" "))  %>% mutate(Count=n())

#getting min age of nodes
df2<-df %>% mutate(node = as.numeric(ifelse(Count > 1, findMRCA(tree, tips=strsplit(Taxa, " ")[[1]]), NA)))
df3<-left_join(df2, t$data, by="node")
df3$Age<-replace_na(df3$Age, 0)


#Summarize min age of nodes
df3 %>% select(locus_name, Age) %>% unique()

annotations<-df3 %>% select(locus_name, node, Age, x, y, Class, virus_genus) %>% unique()

minAges<-annotations %>% rename(MinimumAge=Age) %>% select(virus_genus, locus_name, MinimumAge, Class)

write_csv(minAges, "outputs/105/parvoviridae-min-ages.csv")
```

##Annotating tree
```{r}

nodes<-filter(annotations, node != "NA")
tips<-filter(annotations, Age == 0)
tipSpecies<-df3 %>% filter(locus_name %in% tips$locus_name)

#formatting for carpet plot
carpet<-annotations %>% mutate(NewX=ifelse(Age==0, max(t$data$x), x)) %>% mutate(NewY=ifelse(Age==0, 0, y))
carpet<-carpet %>% mutate(isTip = ifelse(Age==0, "Tip", "Node"))
u<-t + geom_tiplab(size=2) +
  xlim(0,440)+
  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
  geom_tiplab(aes(subset=label %in% tipSpecies$Delim), size=2, color="grey")+
  geom_label_repel(data = nodes, aes(x=x, y=y, label=locus_name), size=2, alpha=0.75)+
  geom_rug(data=filter(carpet, isTip=="Node"), aes(x=NewX, y=NewY, color=virus_genus), sides="b", position="jitter")  +
  theme_tree2()+
  scale_x_continuous(name="Time in Million of Years Ago", breaks=c(35,135,235,335,435),
                     labels=c(400,300,200,100,0), limits=c(0,550))
u
ggsave("outputs/105/parvoviridae-min-ages-timetree.pdf", height=20, width=8.5)

```

##Alternative annotation

```{r}
v<-t + 
  #geom_tiplab(size=2) +
  xlim(0,440)+
#  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
#  geom_tiplab(aes(subset=label %in% tipSpecies$Delim), size=2, color="grey")+
  geom_point(data = nodes, aes(x=x, y=y, fill=virus_genus), size=4, alpha=0.75, pch=23, position = "jitter")+
  #geom_rug(data=filter(carpet, isTip=="Node"), aes(x=NewX, y=NewY, color=virus_genus), sides="b", position="jitter")  +
  theme_tree2()+
  scale_x_continuous(name="\nTime in Million of Years Ago", breaks=c(35,135,235,335,435),
                     labels=c(400,300,200,100,0), limits=c(0,550)) +
  theme(axis.title.x = element_text(size=14)) +
  theme(legend.position = "None")

v

ggsave("outputs/105/parvoviridae-min-ages-timetree-as-nodes.pdf", height=11, width=8.5)

```

Now, I want to add colors to the tips to show the kinds of epvs present. It would be great if it was dotplot like....

```{r}
fact<-10
bigdf<-left_join(data, tdf, by=c("organism"="label")) %>%
  select(Species, organism, virus_genus, x, y, locus_name) %>%
  unique() %>% select(-locus_name) %>% 
  group_by(Species, organism, virus_genus, x, y) %>%
  summarize(Count=n()) %>%
  unique() %>%
  mutate(newX = ifelse(virus_genus=="Amdoparvovirus", x+fact,
                ifelse(virus_genus=="Dependoparvovirus", x+2*fact,
                ifelse(virus_genus=="Erythroparvovirus", x+3*fact,
                ifelse(virus_genus=="Ichthamaparvovirus", x+4*fact,
                ifelse(virus_genus=="Protoparvovirus", x+5*fact, x))))))

```

```{r}
u <-t + 
  #geom_tiplab(size=2) +
  xlim(0,440)+
#  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
#  geom_tiplab(aes(subset=label %in% tipSpecies$Delim), size=2, color="grey")+
  geom_point(data = nodes, aes(x=x, y=y, fill=virus_genus), size=4, alpha=0.75, pch=23, position = "jitter") +
  geom_point(data= bigdf, aes(x=newX, y=y, fill=virus_genus, size=Count), alpha=0.75, pch=21) +
  #geom_rug(data=filter(carpet, isTip=="Node"), aes(x=NewX, y=NewY, color=virus_genus), sides="b", position="jitter")  +
  theme_tree2()+
  scale_x_continuous(name="\nTime in Million of Years Ago", breaks=c(35,135,235,335,435),
                     labels=c(400,300,200,100,0), limits=c(0,550)) +
  theme(axis.title.x = element_text(size=14)) 
u

ggsave("outputs/105/parvoviridae-min-ages-timetree-as-nodes-with-tips.pdf", height=11, width=8.5)

```

How about a dotplot like approach.
stopped working here..
```{r}

#Order is a character string
#Clades is a dataframe "clades" in this markdown
#Tree is our tree, here it is tree
#Adding some classifications to the .tsv

ddd<-read_tsv("outputs/102/epv-dependo-side-data-edited.tsv") %>%
    select(organism, sequenceID, locus_name,host_class,host_order) %>% mutate(virus_genus="Dependoparvovirus") %>%
  mutate(Class=host_class, Order=host_order) %>% select(-host_class, -host_order)

damdo<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-amdo-side-data.tsv") %>% 
  select(organism, virus_genus, sequenceID,Class,Order) %>% mutate(locus_name=paste0("amdo-",sequenceID))


dddf<-rbind(ddd,damdo) %>% filter(locus_name != "NK") %>%
  filter(organism %in% tree$tip.label) %>%
  group_by(locus_name) 

clades<-dddf %>% select(organism, Class, Order)

Clades<-clades
Tree<-tree
species<-ddd %>% select(organism) %>% unique()

findMeANode<-function(Taxon) {
  species <- dplyr::filter(Clades, Order == Taxon)
  speciesNode <- findMRCA(Tree, tips=species$organism)
  return(speciesNode)
}

orders <- clades %>% ungroup %>% select(Order) %>% unique() %>% arrange(Order)

ovector<-orders$Order
findMeANode("Carnivora")
nodes<-lapply(ovector, findMeANode)

nodes[sapply(nodes, is.null)] <- NA

n<-as_tibble(as.vector(unlist(nodes)))

cladelabels<-bind_cols(orders,n) %>% filter(value != "NA")

```

Looks like I have to incldue th eclade labels one at a time.
```{r}
cladelabels<-filter(cladelabels, !(Order %in% c("Unclassified") ))
                    
nn<-cladelabels$value
ll<-cladelabels$Order
v + 
    geom_cladelabel(nn[1],ll[1], offset=15) +
    geom_cladelabel(nn[2],ll[2], offset=15) +
    geom_cladelabel(nn[3],ll[3], offset=15) +
    geom_cladelabel(nn[4],ll[4], offset=15) +
    geom_cladelabel(nn[5],ll[5], offset=15) +
    geom_cladelabel(nn[6],ll[6], offset=15) +
    geom_cladelabel(nn[7],ll[7], offset=15) +
    geom_cladelabel(nn[8],ll[8], offset=15) +
 #   geom_cladelabel(nn[9],ll[9], offset=15) +
    geom_cladelabel(nn[10],ll[10], offset=15) +
    geom_cladelabel(nn[11],ll[11], offset=15) +
    geom_cladelabel(nn[12],ll[12], offset=15) +
    geom_cladelabel(nn[13],ll[13], offset=15) +
    geom_cladelabel(nn[14],ll[14], offset=15) 

ggsave("outputs/102/parvoviridae-min-ages-timetree-as-nodes-clades.pdf", height=11, width=8.5)

```
