---
title: "201-ancestral-again"
author: "Mac Campbell"
date: "5/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Ancestral sequence reconstruction

Let's put these all into outputs/201. First, compute some EVEs that meet our criterion of >= 4 seqs.

```{r}
library(tidyverse)
library(ggtree)
data<-read_tsv("data/epv-side-data.tsv") %>% filter(locus_name !="NK") %>% group_by(locus_name) %>%
  summarize(Count=n()) %>% arrange(-Count) %>% filter(Count >= 4)
data
data %>% arrange(locus_name)
```

Obtain alignments.

```{r}
path<-data %>% mutate(Path=paste0("/Users/mac/github/AAV-GLUE/alignments/eve/tip/",locus_name,".aln.fna")) %>% select(Path)
write_tsv(path, col_names = FALSE, "/Users/mac/github/eve-trees/outputs/201/paths.txt")
```

Only obtained 12 alignments.
There are some differences in naming for loci and *.fna files.   
felid-pinniped vs carnivore
camelid vs camelidae

There are also two entries in the .tsv consulted:
dependo.20-laurasiatheria	9			
dependo.22-laurasiatheria	42			

```{sh, eval=FALSE}
cp /Users/mac/github/AAV-GLUE/alignments/eve/tip/*.fna .
```

In `/Users/mac/github/eve-trees/outputs/201`     
```{sh, eval=FALSE}
#cat paths.txt | while read line; do cp $line .; done;
for f in *.fna; do seqConverter.pl -d$f -ope; done;
for f in *.phylip; do grep -v _con_ $f > `basename $f .phylip`.phy; done;
```

Keep an eye out for a consensus sequence so that the phylips may have too many taxa specified. Yeah.     

### dependo.1
For dependo.1 we can root by the hippo, Hippopotamus_amphibius_Dependo_166_ here.
```{sh, eval=FALSE}
raxmlHPC-PTHREADS-SSE3 -T 3 -m GTRGAMMA -p 123 -f a -# 100 -x 123 -o Hippopotamus_amphibius_Dependo_166_ -n dependo.1 -s dependo.1-whippomorpha.aln.phy

raxmlHPC-PTHREADS-SSE3  -T 3 -m GTRGAMMA -p 123 -f A -t RAxML_bestTree.dependo.1 -s dependo.1-whippomorpha.aln.phy -n dependo.1.ancestral

grep ROOT RAxML_marginalAncestralStates.dependo.1.ancestral | perl -pe 's/ROOT\s/>dependo.1-whippomorpha-ancestral\n/' >dependo.1-whippomorpha.fasta

```

What does it look like?

```{r}
t1<-read.tree("outputs/201/RAxML_nodeLabelledRootedTree.dependo.1.ancestral")
ggtree(t1)+geom_nodelab()+geom_tiplab() + xlim(0,20)
```

### dependo.2
Eptesicus_fuscus_Dependo_120_,Lasiurus_borealis_Dependo_179_,Pipistrellus_pipistrellus_Dependo_295_    
Appear to not be monophyletic in some of the bootstrap trees. What about the "best" tree?    
```{sh, eval=FALSE}
raxmlHPC-PTHREADS-SSE3 -T 3 -m GTRGAMMA -p 123 -f a -# 100 -x 123 -o Eptesicus_fuscus_Dependo_120_,Lasiurus_borealis_Dependo_179_,Pipistrellus_pipistrellus_Dependo_295_ -n dependo.2 -s dependo.2-vespertilionidae.aln.phy

raxmlHPC-PTHREADS-SSE3  -T 3 -m GTRGAMMA -p 123 -f A -t RAxML_bestTree.dependo.2 -s dependo.2-vespertilionidae.aln.phy -n dependo.2.ancestral

grep ROOT RAxML_marginalAncestralStates.dependo.2.ancestral | perl -pe 's/ROOT\s/>dependo.2-vespertilionidae-ancestral\n/' >dependo.2-vespertilionidae.fasta

```

```{r}
t2<-read.tree("outputs/201/RAxML_nodeLabelledRootedTree.dependo.2.ancestral")
ggtree(t2)+geom_nodelab()+geom_tiplab() + xlim(0,12)
```

### dependo.14, 16,17
```{sh, eval=FALSE}
raxmlHPC-PTHREADS-SSE3 -T 3 -m GTRGAMMA -p 123 -f a -# 100 -x 123 -o Cercocebus_atys_Dependo_71_ -n dependo.14 -s dependo.14-cercopithecidae.aln.phy

raxmlHPC-PTHREADS-SSE3  -T 3 -m GTRGAMMA -p 123 -f A -t RAxML_bestTree.dependo.14 -s dependo.14-cercopithecidae.aln.phy -n dependo.14.ancestral

grep ROOT RAxML_marginalAncestralStates.dependo.14.ancestral | perl -pe 's/ROOT\s/>dependo.14-cercopithecidae-ancestral\n/' >dependo.14-cercopithecidae.fasta

raxmlHPC-PTHREADS-SSE3 -T 3 -m GTRGAMMA -p 123 -f a -# 100 -x 123 -o Vulpes_vulpes_Dependo_358_ -n dependo.16 -s dependo.16-canid.aln.phy

raxmlHPC-PTHREADS-SSE3  -T 3 -m GTRGAMMA -p 123 -f A -t RAxML_bestTree.dependo.16 -s dependo.16-canid.aln.phy -n dependo.16.ancestral

grep ROOT RAxML_marginalAncestralStates.dependo.16.ancestral | perl -pe 's/ROOT\s/>dependo.16-canid-ancestral\n/' >dependo.16-canid.fasta

raxmlHPC-PTHREADS-SSE3 -T 3 -m GTRGAMMA -p 123 -f a -# 100 -x 123 -o Panthera_tigris_altaica_Dependo_271_,Puma_concolor_Dependo_321_,Felis_catus_Dependo_148_  -n dependo.17 -s dependo.17-carnivore.aln.phy

raxmlHPC-PTHREADS-SSE3  -T 3 -m GTRGAMMA -p 123 -f A -t RAxML_bestTree.dependo.17 -s dependo.17-carnivore.aln.phy -n dependo.17.ancestral

grep ROOT RAxML_marginalAncestralStates.dependo.17.ancestral | perl -pe 's/ROOT\s/>dependo.17-carnivore-ancestral\n/' >dependo.17-carnivore.fasta
```

```{r}
t14<-read.tree("outputs/201/RAxML_nodeLabelledRootedTree.dependo.14.ancestral")
ggtree(t14)+geom_nodelab()+geom_tiplab() + xlim(0,6)
```

```{r}
t16<-read.tree("outputs/201/RAxML_nodeLabelledRootedTree.dependo.16.ancestral")
ggtree(t16)+geom_nodelab()+geom_tiplab() + xlim(0,6)
```

```{r}
t17<-read.tree("outputs/201/RAxML_nodeLabelledRootedTree.dependo.17.ancestral")
ggtree(t17)+geom_nodelab()+geom_tiplab() + xlim(0,6)
```
### dependo.22

```{sh, eval=FALSE}
raxmlHPC-PTHREADS-SSE3 -T 3 -m GTRGAMMA -p 123 -f a -# 100 -x 123 -o Desmodus_rotundus_Dependo_101_  -n dependo.22 -s dependo.22-laurasiatheria.aln.phy

raxmlHPC-PTHREADS-SSE3  -T 3 -m GTRGAMMA -p 123 -f A -t RAxML_bestTree.dependo.22 -s  dependo.22-laurasiatheria.aln.phy -n dependo.22.ancestral

grep ROOT RAxML_marginalAncestralStates.dependo.22.ancestral | perl -pe 's/ROOT\s/>dependo.22-laurasiatheria-ancestral\n/' > dependo.22-laurasiatheria-ancestral.fasta
```

```{r}
t22<-read.tree("outputs/201/RAxML_nodeLabelledRootedTree.dependo.22.ancestral")
ggtree(t22)+geom_nodelab()+geom_tiplab() + xlim(0,30)
```

### dependo.23

```{sh, eval=FALSE}
raxmlHPC-PTHREADS-SSE3 -T 3 -m GTRGAMMA -p 123 -f a -# 100 -x 123 -o Vicugna_pacos_Dependo_356_   -n dependo.23 -s dependo.23-camelidae.aln.phy

raxmlHPC-PTHREADS-SSE3  -T 3 -m GTRGAMMA -p 123 -f A -t RAxML_bestTree.dependo.23 -s  dependo.23-camelidae.aln.phy -n dependo.23.ancestral

grep ROOT RAxML_marginalAncestralStates.dependo.23.ancestral | perl -pe 's/ROOT\s/>dependo.23-camelid-ancestral\n/' > dependo.23-camelidae-ancestral.fasta
```

```{r}
t23<-read.tree("outputs/201/RAxML_nodeLabelledRootedTree.dependo.23.ancestral")
ggtree(t23)+geom_nodelab()+geom_tiplab() + xlim(0,6)
```

### dependo.26
With more than one sequence in this one, I'm not sure how to root it. Let's take a look.

```{sh, eval=FALSE}
raxmlHPC-PTHREADS-SSE3 -T 3 -m GTRGAMMA -p 123 -f a -# 100 -x 123 -o Daubentonia_madagascariensis_Dependo_95_ -n dependo.26 -s dependo.26-lemuridae.aln.phy
```

```{r}
t26<-read.tree("outputs/201/RAxML_bipartitions.dependo.26")
ggtree(t26)+geom_nodelab()+geom_tiplab() + xlim(0,.5)
```

Propithecus are sister to each other. Maybe it wouldn't be so bad to root by both Dubentonia, though, it looks like there is some support for it to be with (Eulemur+Prolemur).

