---
title: "500-non-dependo-timetrees"
author: "Mac Campbell"
date: "September 29, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ggtree)
library(ggrepel)
```

##Lets see if we have any convincing loci in the Parvoviridae.

###Amdo
Amdo_127319_Ellobius   
Amdo_69149_Ellobius   
Amdo_3304_Ellobius    
Amdo_567_Procavia    
Amdo_565_Orycteropus    

The Ellobius EPV's in part are on the same contig. I think these could be considered the same locus. Probably all independent.

Ellobius not in ENSEMBL.

Let's check out Procavia and Orycteropus....

I don't find evidence that these are two loci, but presumably the two Probothrops are.
Orycteropus afer
Procavia capensis
Ellobius lutescens
Protobothrops flavoviridis
Protobothrops mucrosquamatus

```{r}
data<-read_tsv("outputs/500/meta.tsv")
data
```

```{r}
tree<-read.tree("outputs/500/amdo-taxon-list.nwk")
t<-ggtree(tree)

#Making tip ages = 0 
t$data<-t$data %>% mutate(Age=max(x)-x)

t + geom_tiplab(size=2) +
  geom_treescale()+
  xlim(0,350)+
  geom_nodelab(aes(label=round(Age,2)), hjust=0, size=2)


```
Obtain minimum ages

```{r}
library(phytools)
#Making "Taxa" for tip labs and calculate number of taxa
#Exlude unassigned "NK" locus_names
#filter to tips in tree
treetips<-t$data %>% filter(isTip==TRUE) %>% select(label)

df<-data %>% filter(locus_name != "NK") %>%
  mutate(Delim=gsub(" ","_",Species)) %>%
  filter(Delim %in% treetips$label) %>%
  group_by(locus_name) %>% mutate(Taxa=paste(Delim, collapse=" "))  %>% mutate(Count=n())

#getting min age of nodes
df2<-df %>% mutate(node = as.numeric(ifelse(Count > 1, findMRCA(tree, tips=strsplit(Taxa, " ")[[1]]), NA)))
df3<-left_join(df2, t$data, by="node")
df3$Age<-replace_na(df3$Age, 0)


#Summarize min age of nodes, but including virus_genus
df3 %>% select(locus_name, Age, virus_genus) %>% unique()

annotations<-df3 %>% select(locus_name, virus_genus, node, Age, x, y) %>% unique()
```

Annotate Tree
```{r}

nodes<-filter(annotations, node != "NA")
tips<-filter(annotations, Age == 0)
tipSpecies<-df3 %>% filter(locus_name %in% tips$locus_name)

#formatting for carpet plot
carpet<-annotations %>% mutate(NewX=ifelse(Age==0, max(t$data$x), x)) %>% mutate(NewY=ifelse(Age==0, 0, y))
carpet<-carpet %>% mutate(isTip = ifelse(Age==0, "Tip", "Node"))
u<-t + geom_tiplab(size=2) +
  xlim(0,400)+
  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=2)+
  geom_tiplab(aes(subset=label %in% tipSpecies$Delim), size=2, color="red")+
  geom_label_repel(data = nodes, aes(x=x, y=y, label=locus_name), size=2, alpha=0.75)+
  geom_rug(data=filter(carpet, isTip=="Node"), aes(x=NewX, y=NewY), sides="b", position="jitter")  +
  theme_tree2()+
  scale_x_continuous(name="Time in Million of Years Ago",
                     breaks=c(max(t$data$x)-300,max(t$data$x)-200,max(t$data$x)-100,max(t$data$x)),
                     labels=c(300,200,100,0), limits=c(0,350))

u
```

## Automate

We should be able to make a function for this... Hmm... Or I can hjust redo it all manually.

```{r}
#' @param data metadata dataframe with at least the columns Species, virus_genus, locus_name
#' @param tree path to tree as string e.g. "outputs/500/amdo-taxon-list.nwk"
#' @param genus to make sure we actually get the right label
timetree <- function(data, tree, genus) {

#Make the tree  
  tree<-read.tree(tree)
  t<-ggtree(tree)

  #Making tip ages = 0 
  t$data<-t$data %>% mutate(Age=max(x)-x)
  
#Making "Taxa" for tip labs and calculate number of taxa
#Exlude unassigned "NK" locus_names
#filter to tips in tree
treetips<-t$data %>% filter(isTip==TRUE) %>% select(label)

df<-data %>% filter(locus_name != "NK") %>%
  filter(virus_genus==genus) %>%
  mutate(Delim=gsub(" ","_",Species)) %>%
  filter(Delim %in% treetips$label) %>%
  group_by(locus_name) %>% mutate(Taxa=paste(Delim, collapse=" "))  %>% mutate(Count=n())

#getting min age of nodes
df2<-df %>% mutate(node = as.numeric(ifelse(Count > 1, findMRCA(tree, tips=strsplit(Taxa, " ")[[1]]), NA)))
df3<-left_join(df2, t$data, by="node")
df3$Age<-replace_na(df3$Age, 0)


#Summarize min age of nodes, but including virus_genus
df3 %>% select(locus_name, Age, virus_genus) %>% unique()

annotations<-df3 %>% select(locus_name, virus_genus, node, Age, x, y) %>% unique()

#Annotate and make tree
nodes<-filter(annotations, node != "NA")
tips<-filter(annotations, Age == 0)
tipSpecies<-df3 %>% filter(locus_name %in% tips$locus_name)

#formatting for carpet plot
carpet<-annotations %>% mutate(NewX=ifelse(Age==0, max(t$data$x), x)) %>% mutate(NewY=ifelse(Age==0, 0, y))
carpet<-carpet %>% mutate(isTip = ifelse(Age==0, "Tip", "Node"))
u<-t + geom_tiplab(size=4) +
  xlim(0,400)+
  geom_nodelab(aes(subset=node %in% nodes$node, label=round(Age,2)), hjust=0, size=3)+
  geom_tiplab(aes(subset=label %in% tipSpecies$Delim), size=4, color="red")+
  geom_label_repel(data = nodes, aes(x=x, y=y, label=locus_name), size=3, alpha=0.75)+
  geom_rug(data=filter(carpet, isTip=="Node"), aes(x=NewX, y=NewY), sides="b", position="jitter")  +
  theme_tree2()+
  scale_x_continuous(name="Time in Million of Years Ago",
                     breaks=c(max(t$data$x)-400,
                              max(t$data$x)-300,max(t$data$x)-200,max(t$data$x)-100,max(t$data$x)),
                     labels=c(400,300,200,100,0), limits=c(0,600)) +
  ggtitle(paste0(genus," Timetree")) +
  theme(plot.title = element_text(hjust = 0.5))
return(u)
}
```

# Amdoparvovirus

```{r}
data<-read_tsv("outputs/500/meta.tsv")
atree<-"outputs/500/amdo-taxon-list.nwk"
att<-timetree(data, atree, "Amdoparvovirus")

att
```

# Bocaparvovirus
Only two exammples, Takifugu rubripes and Pteronotus parnelli, these are 433 mya divergent. Probably different loci...... 13 hits on Tru20 with blastn of 180 bp EPV. Searched with tblastx NC_002077.1 and a representative Bocaparvirus, fTakRub1.2 on 20. I'm somewhat confused about this.

```{r}
botree<-"outputs/500/boca-taxon-list.nwk"
bott<-timetree(data, botree, "Bocaparvovirus")
bott
```

# Bufavirus
```{r}
butree<-"outputs/500/bufa-taxon-list.nwk"
butt<-timetree(data, butree, "Bufavirus")
butt
```

# Chapparvovirus
```{r}
cutree<-"outputs/500/chap-taxon-list.nwk"
cutt<-timetree(data, cutree, "Chapparvovirus")
cutt
```

#Erythroparvovirus
Hyaena?

# Print out
```{r}
library(ggpubr)
ggarrange(att, bott, butt, cutt, ncol=2, nrow=2)
ggsave("outputs/500/genera-timetree.pdf", width=11, height=11)
ggsave("outputs/500/genera-timetree.jpg", width=11, height=11)

```

#As facets?
Seems to me, if I treat each tree as a facet instead, we can have the same x-axis scale.

```{r}
library(grid)

grid.draw(rbind(ggplotGrob(att), ggplotGrob(bott)))
```

