---
title: "600-protoparvovirus-timetree"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(plyr)
library(tidyverse)
library(ggtree)
library(cowplot)
```

```{r}
tree<-read.tree("outputs/600/protoparvovirus-host-genera.nwk")
t<-ggtree(tree)

#Making tip ages = 0 
t$data<-t$data %>% mutate(Age=max(x)-x)

t + geom_tiplab(size=2) +
  geom_treescale()+
  xlim(0,200)+
  geom_nodelab(aes(label=round(Age,2)), hjust=0, size=2)


ggsave("outputs/600/protoparvoviridae-timetree.pdf", height=20, width=8.5)
```

##Host ranges

```{r}
library(rgbif)
library(countrycode)
sppToContinents <- function(sp) {
  out<-occ_search(scientificName=sp, limit=20)
  countries<-out$data$country
  if (length(countries[!is.na(countries)]) > 1) {
  continents<-unique(countrycode(sourcevar = countries, origin="country.name",destination="continent"))
  continents<-continents[!is.na(continents)]
  return(cbind(sp, continents)) }
}
```


```{r, eval=FALSE}
species<-read_tsv("outputs/600/protoparvovirus-host-genera.txt", col_names=c("Species"))
spp<-species$Species
clist<-lapply(spp, sppToContinents)

```

```{r, eval=FALSE}
df<-as_tibble( ldply(clist, data.frame))
save(df,file="outputs/600/df.rda")

df
write_csv(df, "outputs/600/species-ranges.csv")
```

Needs a bit of help.
```{r}
rev<-read_csv("outputs/600/species-ranges-curated.csv")
rev
```

#Now to set up DEC

datafile looks like:   
NumTips NumStates     
SpeciesName 010010    

etc. Seven States:
North America
South America
Caribbean
Europe
North Africa
Asia
Australia     
"protoData.txt"      

protoConfigure.txt        
need a file like:      

treefile = prunedTree.tre     
datafile = abudefduf.data     
areanames = WPac CPac EPac WAtl EAtl WInd EInd     
ancstate = _all_   
states     

need _ in all names to match tree....      
timetree.org tree has node numbers in there, removing...      
Typo in ancstate.....
Now it works!!!
```{sh, eval=FALSE}

~/lagrange/lagrange_cpp_0.1BETA2_for_MAC10.6/lagrange_cpp protoConfigure.txt > output

```
     
## Plot
```{r}

tree<-read.tree(file="outputs/600/protoexport.tre.bgstates.tre")
nodeLabels<-tree$node.label

##Modified by Mac for making labels
##' add node label layer
##'
##'
##' @title geom_nodelab
##' @param mapping aes mapping
##' @param nudge_x horizontal adjustment to nudge label
##' @param nudge_y vertical adjustment to nudge label
##' @param geom one of 'text', 'label', 'image' and 'phylopic'
##' @param hjust horizontal alignment, one of 0, 0.5 or 1
##' @param ... additional parameters
##' @return geom layer
##' @export
##' @author guangchuang yu
geom_nodelab <- function(mapping = NULL, nudge_x = 0, nudge_y = 0, geom = "label", hjust = 0.5, ...) {
  self_mapping <- aes_(subset = ~!isTip)
  if (is.null(mapping)) {
    mapping <- self_mapping
  } else {
    mapping <- modifyList(self_mapping, mapping)
  }
  
  geom_tiplab(mapping, offset = nudge_x, nudge_y = nudge_y, geom = geom, hjust = hjust, ...)
}

#geom_nodelab()
p<-ggtree(tree, right=TRUE)+geom_nodelab(fill="grey50", color="white", size=2.5)+
  geom_tiplab(size=3)+
  theme_tree2()+
  ggplot2::xlim(-200, 50)


p$data$x <- p$data$x - max(p$data$x)

p

ggsave("outputs/600/protoparvovirus-DEC-tree.pdf")
```

Might be nice to have node ages as a plot.

```{r}
q<-ggtree(tree, right=TRUE)

q$data$x <- q$data$x - max(q$data$x)
q$data<-q$data %>% mutate(label=ifelse(isTip=="TRUE",label,round(x,2)))
 
q2<-q + geom_nodelab(aes(label=label), color="black", size=2.5)+
  geom_tiplab(size=3)+
  theme_tree2()+
  ggplot2::xlim(-200, 50)

q2

```


All together now!    
```{r}
plot_grid(p, q2, ncol=2)
ggsave("outputs/600/protoparvoviridae-timetree-with-ancdists.pdf", width = 11, height=8.5)
```