---
title: "601-dependoparvovirus"
author: "Mac Campbell"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(plyr)
library(tidyverse)
library(ggtree)
library(cowplot)
```

#Get hosts
```{r}
dependo<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-dependo-side-data.tsv") %>%
    select(organism, sequenceID, locus_name,host_class) %>% mutate(virus_genus="Dependoparvovirus") %>%
  mutate(Class=host_class) %>% select(-host_class)

data<-dependo
data$Species<-gsub("_"," ",data$organism)
data<-data %>% select(Species, organism, locus_name, virus_genus, sequenceID, Class)
species<-data %>% select(Species) %>% unique()
write_tsv(species, "outputs/601/dependo-species-list.tsv", col_names = FALSE)

```

Send to timetree.....
17 naming issues
Equus caballus fjord
Equus caballus icelandic
Panthera tigris altaica (replaced with Panthera tigris)
Canis familiaris (replaced with Canis lupus)
Canis lupus dingo (replaced with Canis lupus)
Equus asinus somalicus (replaced with Equus asinus)
Equus africanus asinus (replaced with Equus asinus)
Equus burchellii boehmi (replaced with Equus burchellii)
Equus zebra hartmannae (replaced with Equus zebra)
Ceratotherium simum cottoni (replaced with Ceratotherium simum)
Dicerorhinus sumatrensis sumatrensis (replaced with Dicerorhinus sumatrensis)
Cervus elaphus hippelaphus (replaced with Cervus elaphus)
Murina aurata feae (replaced with Murina aurata)
Mus castaneus (replaced with Mus musculus)
Peromyscus maniculatus bairdii (replaced with Peromyscus maniculatus)
Colobus angolensis palliatus (replaced with Colobus angolensis)
Neophocaena asiaeorientalis (no substitute found)

```{r}
tree<-read.tree("outputs/601/dependo-species-list.nwk")
t<-ggtree(tree)

#Making tip ages = 0 
t$data<-t$data %>% mutate(Age=max(x)-x)

t + geom_tiplab(size=2) +
  geom_treescale()+
  xlim(0,200)+
  geom_nodelab(aes(label=round(Age,2)), hjust=0, size=2)


ggsave("outputs/601/dependoparvovirus-timetree.pdf", height=20, width=8.5)
```

##Host ranges

```{r}
library(rgbif)
library(countrycode)
sppToContinents <- function(sp) {
  out<-occ_search(scientificName=sp, limit=50)
  countries<-out$data$country
  if (length(countries[!is.na(countries)]) > 1) {
  continents<-unique(countrycode(sourcevar = countries, origin="country.name",destination="continent"))
  continents<-continents[!is.na(continents)]
  return(cbind(sp, continents)) }
}
```


```{r, eval=FALSE}
species<-read_tsv("outputs/601/dependo-species-list.tsv", col_names=c("Species"))
spp<-species$Species
clist<-lapply(spp, sppToContinents)

```

```{r, eval=FALSE}
df<-as_tibble( ldply(clist, data.frame))
save(df,file="outputs/601/df.rda")

df
write_csv(df, "outputs/601/species-ranges.csv")
```

Needs a bit of help.
```{r}
rev<-read_csv("outputs/601/species-ranges-curated.csv")
rev
```

#Now to set up DEC

datafile looks like:   
NumTips NumStates     
SpeciesName 010010    

etc. Ten States:
North America1
South America2
Caribbean3
Europe4
North Africa5
Sub-Saharan Africa6    
Asia7     
Australia8     
Madagascar9
Antartica10
"dependoData.txt"      

dependoConfigure.txt         
need a file like:      

treefile = prunedTree.tre     
datafile = abudefduf.data     
areanames = WPac CPac EPac WAtl EAtl WInd EInd     
ancstate = _all_   
states     

```{sh, eval=FALSE}

~/lagrange/lagrange_cpp_0.1BETA2_for_MAC10.6/lagrange_cpp dependoConfigure.txt > output

```
     
## Plot
```{r}

tree<-read.tree(file="outputs/601/tree.nwk.bgstates.tre")
nodeLabels<-tree$node.label

##Modified by Mac for making labels
##' add node label layer
##'
##'
##' @title geom_nodelab
##' @param mapping aes mapping
##' @param nudge_x horizontal adjustment to nudge label
##' @param nudge_y vertical adjustment to nudge label
##' @param geom one of 'text', 'label', 'image' and 'phylopic'
##' @param hjust horizontal alignment, one of 0, 0.5 or 1
##' @param ... additional parameters
##' @return geom layer
##' @export
##' @author guangchuang yu
geom_nodelab <- function(mapping = NULL, nudge_x = 0, nudge_y = 0, geom = "label", hjust = 0.5, ...) {
  self_mapping <- aes_(subset = ~!isTip)
  if (is.null(mapping)) {
    mapping <- self_mapping
  } else {
    mapping <- modifyList(self_mapping, mapping)
  }
  
  geom_tiplab(mapping, offset = nudge_x, nudge_y = nudge_y, geom = geom, hjust = hjust, ...)
}

#geom_nodelab()
p<-ggtree(tree, right=TRUE)+geom_nodelab(fill="grey50", color="white", size=2.5)+
  geom_tiplab(size=3)+
  theme_tree2()+
  ggplot2::xlim(-200, 50)


p$data$x <- p$data$x - max(p$data$x)

p

ggsave("outputs/601/dependoparvovirus-DEC-tree.pdf")
```

Might be nice to have node ages as a plot.

```{r}
q<-ggtree(tree, right=TRUE)

q$data$x <- q$data$x - max(q$data$x)
q$data<-q$data %>% mutate(label=ifelse(isTip=="TRUE",label,round(x,2)))
 
q2<-q + geom_nodelab(aes(label=label), color="black", size=2.5)+
  geom_tiplab(size=3)+
  theme_tree2()+
  ggplot2::xlim(-200, 50)

q2

```


All together now!    
```{r}
plot_grid(p, q2, ncol=2)
ggsave("outputs/601/dependoparvovirus-timetree-with-ancdists.pdf", width = 11, height=8.5)
```