---
title: "800-ancestral-dists"
author: "Mac Campbell"
date: "12/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ggtree)
library(ape)
library(paleotree)
library(phytools)
```

## Ancestral ranges for taxa
Springer et al. (2011) - https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3138613/      
Laurasiatheria - Eurasia     
Boreoeutheria - Eurasia   
Euarchontoglires - Eurasia   
Africa - Afrotheria
Xenartha -  South America

## Let's take a look at the tree:
```{r}
rep<-read.tree("~/github/Parvoviridae-GLUE/trees/genus-level-with-eves/dependo-root-rep78.eve.export_nucs.tre")
repmeta<-read_tsv("~/github/Parvoviridae-GLUE/trees/genus-level-with-eves/dependo-root-rep78.eve.figtree-annotations.tsv")
t <-  ggtree(rep) %<+% repmeta+geom_tiplab(aes(label=sequenceID)) + 
    xlim(0,5)

t
```

## Make a time-scaled tree with paleotree
Setting all tip dates to 0 for non-eves, can left_join on ages from elsewhere.
```{r}
ages<-repmeta %>% select(taxon, sequenceID, name) %>% mutate(locus_name=sequenceID)
ages$locus_name<-gsub("-con","",ages$locus_name)
mins<-read_csv("outputs/102/parvoviridae-min-ages.csv")

ages2<-ages %>% left_join(mins)
```

Not complete agreement between *.tsv used to generate min ages and this tree.

```{r}
ages3 <- ages2 %>% mutate(MinimumAge = ifelse(startsWith(sequenceID,"dependo")=="FALSE", 0, MinimumAge))
#remove those with no info
ages4<-ages3[is.na(ages3$MinimumAge)=="FALSE",]
drops<-ages3[is.na(ages3$MinimumAge)=="FALSE",]

tree2<-keep.tip(rep, tip=ages4$taxon)
```

#### What's our tree look like?
```{r}
t2<- ggtree(midpoint.root(tree2)) %<+% ages4 + geom_tiplab(aes(label=sequenceID)) + 
    xlim(0,5)

t2
```

Now we should be able to make it time calibrated....         
tree<-unscaled cladogram       
timeData<-two-column matrix of first and last occurrences in absolute continuous time.     
```{r}
#reorder data to rename tips
order<-as_tibble(tree2$tip.label) %>% left_join(ages4, by=c("value"="taxon")) %>% mutate(Age2=MinimumAge)
tree3<-midpoint.root(tree2)
tree3$tip.label<-order$sequenceID

matrix<-order %>% select(MinimumAge, Age2) %>% as.matrix()
row.names(matrix)<-tree3$tip.label
## Root by oldest clade
ggtree(tree3) + geom_tiplab()+geom_nodelab(aes(label=node))
tree4<-root(tree3, node = 65)
ggtree(tree4) + geom_tiplab()+geom_nodelab(aes(label=node))
```

```{r}
pdf("outputs/800/paleo-tree.pdf", height=14, width=8)
tt<-timePaleoPhy(tree4,matrix, ntrees = 1, plot = TRUE)
dev.off()
```

ggtree versions

```{r}
max<-tt$root.time

ggtree(tt) + 
  geom_tiplab() + 
  theme_tree2() +
  scale_x_continuous(name="\nTime in Million of Years Ago", breaks=c(max-80,max-60,max-40,max-20,max),
                     labels=c(80,60,40,20,0), limits=c(0,120)) +
  theme(axis.title.x = element_text(size=14, face="bold")) +
  theme(axis.text.x = element_text(size=10, face="bold"))
ggsave("outputs/800/paleo-tree-ggtree.pdf", height=7, width=8)
```

