---
title: "901-dependo-tree-with-sizes"
author: "Mac Campbell"
date: "February 5, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Dependo Tree with Clade Sizes

```{r}
library(tidyverse)
library(ggtree)
library(ape)
library(phytools)
```

```{r}
cap<-read.tree("~/github/Parvoviridae-GLUE/trees/genus-level-with-eves/dependo-root-vp1.eve.export_nucs.tre")
cap<-midpoint.root(cap)
capmeta<-read_tsv("~/github/Parvoviridae-GLUE/trees/genus-level-with-eves/dependo-root-vp1.eve.figtree-annotations.tsv")
t2<-ggtree(cap) %<+% capmeta
```

```{r}
t2 + geom_tiplab(aes(label=sequenceID), cex=2) + xlim(0,4) 
```


Now to figure out the size of each locus.

```{r}
dependo<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-dependo-side-data.tsv") %>%
    select(organism, sequenceID, locus_name,host_class,host_superorder) %>% mutate(virus_genus="Dependoparvovirus") %>%
  mutate(Class=host_class,`Host Superorder`=host_superorder) %>% select(-host_class,-host_superorder)
```

```{r}
counts<-dependo %>% group_by(Class, `Host Superorder`,locus_name) %>% summarize(Count=n()) %>% rename(sequenceID=locus_name)
```

Some of the meta ends with con...

```{r}
t3<-t2
t3$data$sequenceID<-gsub("-con","",t3$data$sequenceID)
t3$data<-left_join(t3$data, counts)
```

```{r}
t3 + geom_tiplab(aes(label=sequenceID), cex=2, align = TRUE) + xlim(0,4) +
  geom_tippoint(aes(size=Count, fill=`Host Superorder`), x=2.75, pch=21, alpha=0.75)
ggsave("outputs/901/dependo-tree-with-counts.pdf")
```

##Example from phyltools
Need something like this:    
trans<-data.frame(tip.label,clade.label,N,depth)       
  tip.label clade.label  N     depth    
1        t3           A 11 0.5209987    
2        t7           B  3 0.4694350    
3        t2           C 17 0.5515002   


```{r}

tips<-filter(t3$data, isTip==TRUE) %>% select(label, name, Count)
tips$Count<-replace_na(tips$Count, 1)
tips<-tips %>% rename(tip.label=label, clade.label=name, N=Count)

tree<-cap
tip.label<-tree$tip.label

depth<-sapply(tip.label,function(x,y) 
	0.5*y$edge.length[which(tree$edge[,2]==
	which(y$tip.label==x))],y=tree)

depth<-as.data.frame(depth)

tips$depth<-depth$depth
obj<-phylo.toBackbone(tree,tips)
plot(obj)

pdf("outputs/901/dependo-backbone-phylo.pdf", width=8, height=11)
plot(obj, cex=.6)
dev.off()
```

##What about for the other gene?

```{r}
cap<-read.tree("~/github/Parvoviridae-GLUE/trees/genus-level-with-eves/dependo-root-rep78.eve.export_nucs.tre")
cap<-midpoint.root(cap)
capmeta<-read_tsv("~/github/Parvoviridae-GLUE/trees/genus-level-with-eves/dependo-root-rep78.eve.figtree-annotations.tsv")
t2<-ggtree(cap) %<+% capmeta
```


```{r}
t2 + geom_tiplab(aes(label=sequenceID), cex=2) + xlim(0,4) 
```


Now to figure out the size of each locus.

```{r}
dependo<-read_tsv("~/github/Parvoviridae-GLUE/tabular/eve/epv-dependo-side-data.tsv") %>%
    select(organism, sequenceID, locus_name,host_class,host_superorder) %>% mutate(virus_genus="Dependoparvovirus") %>%
  mutate(Class=host_class,`Host Superorder`=host_superorder) %>% select(-host_class,-host_superorder)
```

```{r}
counts<-dependo %>% group_by(Class, `Host Superorder`,locus_name) %>% summarize(Count=n()) %>% rename(sequenceID=locus_name)
```

Some of the meta ends with con...

```{r}
t3<-t2
t3$data$sequenceID<-gsub("-con","",t3$data$sequenceID)
t3$data<-left_join(t3$data, counts)
```

```{r}
t3 + geom_tiplab(aes(label=sequenceID), cex=2, align = TRUE) + xlim(0,4) +
  geom_tippoint(aes(size=Count, fill=`Host Superorder`), x=2.75, pch=21, alpha=0.75)
ggsave("outputs/901/dependo-tree-rep78-with-counts.pdf")
```

##Example from phyltools
Need something like this:    
trans<-data.frame(tip.label,clade.label,N,depth)       
  tip.label clade.label  N     depth    
1        t3           A 11 0.5209987    
2        t7           B  3 0.4694350    
3        t2           C 17 0.5515002   


```{r}

tips<-filter(t3$data, isTip==TRUE) %>% select(label, name, Count)
tips$Count<-replace_na(tips$Count, 1)
tips<-tips %>% rename(tip.label=label, clade.label=name, N=Count)

tree<-cap
tip.label<-tree$tip.label

depth<-sapply(tip.label,function(x,y) 
	0.5*y$edge.length[which(tree$edge[,2]==
	which(y$tip.label==x))],y=tree)

depth<-as.data.frame(depth)

tips$depth<-depth$depth
obj<-phylo.toBackbone(tree,tips)
plot(obj)

pdf("outputs/901/dependo-backbone-phylo-rep78.pdf", width=8, height=15)
plot(obj, cex=.6)
dev.off()
```
